---
title: "Summary statistics"
author: "Habtu et al."
date: "2026-02-25"
output: html_document
---

#One-off
setwd("~/slimwork/New/Latest/One_off_V1/Null/Female/NB2/low_chancest3")**Trial20NBRevised**
setwd("~/slimwork/New/Latest/One_off_V1/Null/Female/NB2/low_chancest2")**Trial20NBRevised_** #Do_nothing
setwd("~/slimwork/New/Latest/One_off_V2/Ten/Female/benefit_two/low_chancest2")**MAM_Fem_B2y1Revised_**
setwd("~/slimwork/New/Latest/One_off_V3edited/Null/Female/benefit_three/low_chancest2")**Trial5_BarbaraallB3nRevised_**
#Ongoing
setwd("~/slimwork/New/Latest/new2Ongoing/Two/Female/benefit_one/low_chancest2")
setwd("~/slimwork/New/Latest/Ongoing_V3L1/Fourteen/Female/benefit_three/high_chancestext")**Trial_B3_Neutral_Females_**
setwd("~/slimwork/New/Latest/newOngoing/Two/Female/benefit_two/low_chancest2")**Trial_Revised_new**
setwd("~/slimwork/New/Latest/newOngoing/Two/Female/NB/low_chancest2") **Trial_allOn0NBRevised_**

```{r}
 # setwd("~/slimwork/New/Latest/newOngoing/Twelve/Female/NB/low_chancest3") 
#setwd("~/slimwork/New/Latest/One_off_V1/Null/Female/NB2/low_chancest2")
setwd("~/slimwork/New/Latest/One_off_V3edited/Ten/Female/benefit_three/low_chancest2")
reps = 1000
popSizeWP <- vector(mode='list', length=reps)
PropDisWP<- vector(mode='list', length=reps)
NeutHetWP<- vector(mode='list', length=reps)
PropBenWP<- vector(mode='list', length=reps)
NumBenWP<- vector(mode='list', length=reps)

for (i in 1:reps) {
  outfileName <- paste("Trial5_BarbaraallB3nRevised_", i, ".csv", sep="")
  dataf <- read.csv (outfileName, skip=17, header=F)
  colnames(dataf) <- c("sim.cycle" ," WPS","mean_fitness","Funhet2" , "Heterozygosity2", "WPMaxAge", "IPMaxAge","IPS", "Heterozygosity","AveageIP","Suscepl","Infectedl","infected_Y1Tot","infected_Y2Tot","infected_Y3Tot","Prop_Inf","Prop_Ben","D_fitness","ND_fitness","infected_Ben_fitness"," uninfected_Ben_fitness","infected_noBen_fitness","uninfected_noBen_fitness","Inbreeding","Inbreedingneut","Inbreeding2","Inbreeding2fun","Inbreeding2neut","males","females","Num_Ben","Num_Ben_Individuals","Prop_Bencheck")

  
  popSizeWP[[i]] <- dataf[,2]
  PropDisWP[[i]] <- dataf[,16]
  NeutHetWP[[i]] <- dataf[,5]
  PropBenWP[[i]] <- dataf[,33]
  NumBenWP[[i]] <- dataf[,32]
}
maxGen <- max(sapply(popSizeWP, length))
popMat <- matrix(0, maxGen, reps)
popMat2 <- matrix(0, maxGen, reps)
popMat3 <- matrix(0, maxGen, reps)
popMat4 <- matrix(0, maxGen, reps)
popMat5 <- matrix(0, maxGen, reps)
for (i in 1:reps){
  thisLen <- length(popSizeWP[[i]])
  thisLen2 <- length(PropDisWP[[i]])
  thisLen3 <- length(PropBenWP[[i]])
  thisLen4 <- length(NeutHetWP[[i]])
  thisLen5 <- length(NumBenWP[[i]])
  popMat[1:thisLen, i] <- popSizeWP[[i]]
  popMat2[1:thisLen2, i] <- PropDisWP[[i]]
  popMat3[1:thisLen3, i] <- PropBenWP[[i]]
  popMat4[1:thisLen4, i] <- NeutHetWP[[i]]
  popMat5[1:thisLen4, i] <- NumBenWP[[i]]
}

B1_Low1= as.data.frame(popMat)
B1_Low3= as.data.frame(popMat2)
B1_Low5= as.data.frame(popMat3)
B1_Low7= as.data.frame(popMat4)
B1_Low9= as.data.frame(popMat5)

ncols=10

satisfying_columns <- which(B1_Low5[100, ] > 0.5)
size <- min(3, length(satisfying_columns))  # At least 3 columns, but not more than available
sampled_satisfying_columns <- sample(satisfying_columns, size)
remaining_columns <- setdiff(seq_len(ncol(B1_Low1)), sampled_satisfying_columns)
sampled_remaining_columns <- sample(remaining_columns, ncols - length(sampled_satisfying_columns))
sampled_columns <- c(sampled_satisfying_columns, sampled_remaining_columns)


#sampled_columns <- sample(ncol(B1_Low1), ncols, replace = FALSE)


Years<-1:1000
#sampled_columns<-1:701
B1_Low2= B1_Low1[Years,sampled_columns]
B1_Low4= B1_Low3[Years,sampled_columns]
B1_Low6= B1_Low5[Years,sampled_columns]
B1_Low8= B1_Low7[Years,sampled_columns]
B1_Low10= B1_Low9[Years,sampled_columns]
# Transform to long format
long_data <- B1_Low2%>%
  mutate(Year = row_number()) %>%
  pivot_longer(cols = starts_with("V"), names_to = "Rep", values_to = "PopSize") %>%
  mutate(
    Year = as.factor(Year),
    Rep = as.integer(str_replace(Rep, "V", ""))
  ) %>%
  arrange(Rep, as.integer(Year)) %>%
  dplyr:: mutate(Rep = as.factor(Rep))

long_data2 <- B1_Low4%>%
  mutate(Year = row_number()) %>%
  pivot_longer(cols = starts_with("V"), names_to = "Rep", values_to = "PropInf") %>%
  mutate(
    Year = as.factor(Year),
    Rep = as.integer(str_replace(Rep, "V", ""))
  ) %>%
  arrange(Rep, as.integer(Year)) %>%
  mutate(Rep = as.factor(Rep))%>%
  dplyr:: select(PropInf)

# Transform to long format
long_data3 <- B1_Low6%>%
  mutate(Year = row_number()) %>%
  pivot_longer(cols = starts_with("V"), names_to = "Rep", values_to = "Prop_Bencheck") %>%
  mutate(
    Year = as.factor(Year),
    Rep = as.integer(str_replace(Rep, "V", ""))
  ) %>%
  arrange(Rep, as.integer(Year)) %>%
  mutate(Rep = as.factor(Rep))%>%
  dplyr:: select(Prop_Bencheck)

long_data4 <- B1_Low8%>%
  mutate(Year = row_number()) %>%
  pivot_longer(cols = starts_with("V"), names_to = "Rep", values_to = "NeutHet") %>%
  mutate(
    Year = as.factor(Year),
    Rep = as.integer(str_replace(Rep, "V", ""))
  ) %>%
  arrange(Rep, as.integer(Year)) %>%
  mutate(Rep = as.factor(Rep))%>%
  dplyr:: select(NeutHet)

long_data5 <- B1_Low10%>%
  mutate(Year = row_number()) %>%
  pivot_longer(cols = starts_with("V"), names_to = "Rep", values_to = "NumBenWP") %>%
  mutate(
    Year = as.factor(Year),
    Rep = as.integer(str_replace(Rep, "V", ""))
  ) %>%
  arrange(Rep, as.integer(Year)) %>%
  mutate(Rep = as.factor(Rep))%>%
  dplyr:: select(NumBenWP)

```


```{r}
setwd("~/slimwork/Plots")
d <- data.frame(long_data, long_data2, long_data3, long_data4, long_data5)
nreps <- 1000

# Determine the maximum available year in popMat
max_available_row <- nrow(popMat)
max_year <- max_available_row  # Assuming row number corresponds to year

# If year 700 is requested but not available, use the maximum available instead
target_year <- 700
analysis_row <- min(target_year, max_available_row)

# Calculate extinction proportion based on available data
extinct_Reps <- sum(popMat[analysis_row, ] == 0, na.rm = TRUE)
Prop.extinction <- round((extinct_Reps) / nreps, 2)
cat(paste("Proportion of simulations went extinct at year", analysis_row, ":", Prop.extinction, "\n"))

# Function to create consistent plot with different color schemes
create_plot <- function(data, color_var, color_title, color_option, extinction_prop, max_year) {
  # Filter data to only include available years if needed
  filtered_data <- data
  
  # Determine x-axis breaks based on available years
  available_levels <- levels(filtered_data$Year)
  x_breaks <- available_levels[seq(1, length(available_levels), by = 10)]
  
  # Create the plot
  p <- ggplot() + 
    geom_line(data = filtered_data, aes(x = Year, y = PopSize, group = Rep, color = !!sym(color_var)), alpha = 0.4) +
    scale_y_continuous(limits = c(0, NA)) +
    scale_x_discrete(breaks = x_breaks) +
    scale_color_viridis_c(
      option = color_option, limits = c(0, 1), begin = 0, end = 0.5
    ) +
    labs(color = color_title, y = "Population size", x = "Year") +
    geom_point()
  
  # Find a position for the text that's guaranteed to be within the plot area
  x_idx <- which(as.numeric(available_levels) >= 70)[1]
  if (is.na(x_idx)) {
    x_idx <- length(available_levels) / 2  # Use middle of x-axis if 70 not available
  }
  x_val <- available_levels[x_idx]
  
  # Add extinction text as a separate annotation
  p <- p + annotate(
    "text",
    x = x_val,
    y = 15,
    label = paste("Proportion of extinction =", Prop.extinction),
    size = 4.5
  )
  
  # Only add vertical lines if those years exist in the data
  if ("2" %in% available_levels) {
    p <- p + geom_vline(xintercept = "2", linetype = "dashed", color = "green", linewidth = 0.8)
  }
  
  if ("6" %in% available_levels) {
    p <- p + geom_vline(xintercept = "6", linetype = "dashed", color = "black", linewidth = 0.8)
  }
  
  # Add theme elements
  p <- p + theme_minimal() +
    theme(
      axis.line.x = element_line(color = "black", linewidth = 0.6),
      axis.line.y.left = element_line(color = "black", linewidth = 0.6),
      axis.ticks = element_line(),
      axis.text.x = element_text(size = 14),        
      axis.text.y = element_text(size = 14),        
      axis.title.x = element_text(size = 16),      
      axis.title.y = element_text(size = 16),      
      legend.position = "right",
      panel.grid.major = element_blank(),    
      panel.grid.minor = element_blank(),
      panel.background = element_rect(fill = "white", color = NA),
      plot.background = element_rect(fill = "white", color = NA)
    ) +
    guides(color = guide_colorbar(title = paste0("Percent ", sub("Prop", "", color_title)), barwidth = 1, barheight = 15))
  
  return(p)
}

# Create the first plot with Prop_Bencheck
plot1 <- create_plot(
  data = d,
  color_var = "Prop_Bencheck", 
  color_title = "Beneficial", 
  color_option = "viridis",
  extinction_prop = Prop.extinction,
  max_year = max_year
)

# Save the first plot
plot_filename1 <- paste0("Trial_Ong_NB_14_P_0.05_R_PropBen.png")
ggsave(plot_filename1, plot = plot1, device = ragg::agg_png, width = 8, height = 7, dpi = 300)

# Create the second plot with PropInf
plot2 <- create_plot(
  data = d,
  color_var = "PropInf", 
  color_title = "Diseased", 
  color_option = "plasma",
  extinction_prop = Prop.extinction,
  max_year = max_year
)

# Save the second plot
plot_filename2 <- paste0("Trial_Ong_NB_14_P_0.05_R_R_WPS.png")
ggsave(plot_filename2, plot = plot2, device = ragg::agg_png, width = 8, height = 7, dpi = 300)

# Print information about what was done
cat(paste("Analysis performed using data up to year/row", analysis_row, "(max available:", max_available_row, ")\n"))
cat(paste("Plots saved as", plot_filename1, "and", plot_filename2, "\n"))

```



```{r}
nreps=1000

# Function to safely extract values from a matrix at a specific row
# Returns a vector of zeros if row is out of bounds
safe_extract_row <- function(matrix, row_index) {
  if(is.null(matrix) || nrow(matrix) < row_index || row_index <= 0) {
    return(rep(NA, ncol(matrix)))
  } else {
    return(matrix[row_index, ])
  }
}

# Function to safely calculate mean and SD with proper handling
safe_mean_sd <- function(values) {
  if(all(values == 0)) {
    return(list(mean = 0.00, sd = 0.00))
  } else {
    return(list(mean = round(mean(values, na.rm = TRUE), 2), 
                sd = round(sd(values, na.rm = TRUE), 2)))
  }
}

# Function to safely calculate extinction
safe_extinction <- function(matrix, row_index, nreps) {
  if(is.null(matrix) || nrow(matrix) < row_index || row_index <= 0) {
    return(100)
  } else {
    extinct_reps <- sum(matrix[row_index, ] == 0, na.rm = TRUE)
    return(((extinct_reps)/nreps) * 100)
  }
}

# Function to safely calculate proportion with BM
safe_proportion_BM <- function(popMat, popMat3, row_index) {
  if(is.null(popMat) || is.null(popMat3) || 
     nrow(popMat) < row_index || nrow(popMat3) < row_index || 
     row_index <= 0) {
    return(0.00)
  } else {
    indices <- which(popMat3[row_index, ] > 0)
    if(length(indices) == 0) {
      return(0.00)
    }
    satisfying_elements <- popMat[row_index, indices]
    satisfying_vector <- as.vector(satisfying_elements)
    proportion <- round((length(satisfying_vector)/length(as.vector(popMat[row_index, ]))) * 100, 2)
    return(proportion)
  }
}

# Check and extract Population Size for Year 50
AveragePS50 <- safe_extract_row(popMat, 51)
PS50_stats <- safe_mean_sd(AveragePS50)
MeanPS50 <- PS50_stats$mean
SDPS50 <- PS50_stats$sd
cat("Mean PS In Year 50:", MeanPS50, "\n")
cat("SD PS In Year 50:", SDPS50, "\n")

# Check and extract Population Size for Year 700
AveragePS700 <- safe_extract_row(popMat, 700)
PS700_stats <- safe_mean_sd(AveragePS700)
MeanPS700 <- PS700_stats$mean
SDPS700 <- PS700_stats$sd
cat("Mean PS In Year 700:", MeanPS700, "\n")
cat("SD PS In Year 700:", SDPS700, "\n")

# Check and extract Heterozygosity for Year 50
AverageHet50 <- safe_extract_row(popMat4, 51)
Het50_stats <- list(
  mean = if(all(AverageHet50 == 0)) 0.00 else mean(AverageHet50, na.rm = TRUE),
  sd = if(all(AverageHet50 == 0)) 0.00 else sd(AverageHet50, na.rm = TRUE)
)
MeanHet50 <- Het50_stats$mean
SDHet50 <- Het50_stats$sd
cat("Mean Het In Year 50:", MeanHet50, "\n")
cat("SD Het In Year 50:", SDHet50, "\n")

# Check and extract Heterozygosity for Year 700
AverageHet700 <- safe_extract_row(popMat4, 700)
Het700_stats <- list(
  mean = if(all(AverageHet700 == 0)) 0.00 else mean(AverageHet700, na.rm = TRUE),
  sd = if(all(AverageHet700 == 0)) 0.00 else sd(AverageHet700, na.rm = TRUE)
)
MeanHet700 <- Het700_stats$mean
SDHet700 <- Het700_stats$sd
cat("Mean Het In Year 700:", MeanHet700, "\n")
cat("SD Het In Year 700:", SDHet700, "\n")

# Calculate Proportion of Extinction for Year 50
Prop.extinction50 <- safe_extinction(popMat4, 51, nreps)
cat("Proportion of simulations went extinct in year 50:", Prop.extinction50, "\n")

# Calculate Proportion of Extinction for Year 700
Prop.extinction700 <- safe_extinction(popMat, 600, nreps)
cat("Proportion of simulations went extinct in year 700:", Prop.extinction700, "\n")

# Calculate Proportion of individuals with BM at year 700
# Note: Original code was using popMat3[600, ], keeping the same row index
Proportion_of_BM <- safe_proportion_BM(popMat, popMat3, 600)
cat("PFM700:", Proportion_of_BM, "\n")

# Calculate Proportion of individuals with BM at year 50
Proportion_of_BM2 <- safe_proportion_BM(popMat, popMat3, 51)
cat("PFM50:", Proportion_of_BM2, "\n")

# Create a data frame for the summary table with parameters as columns and values as rows
summary_table <- data.frame(
  PS_50 = sprintf("%.2f (%.2f)", MeanPS50, SDPS50),
  PS_700 = sprintf("%.2f (%.2f)", MeanPS700, SDPS700),
  Het_50 = sprintf("%.2e (%.2e)", MeanHet50, SDHet50),
  Het_700 = sprintf("%.2e (%.2e)", MeanHet700, SDHet700),
  Extinction_50 = Prop.extinction50,
  Extinction_700 = Prop.extinction700,
  PBM_50 = Proportion_of_BM2,
  PBM_700 = Proportion_of_BM
)

# Display the summary table with HTML formatting
kable(summary_table, 
      format = "latex", 
      row.names = TRUE,  # Adds row indices for easier reading
      caption = "Summary of Simulation Results") %>%
  kable_styling(full_width = FALSE, position = "left")

# Display the summary table with LaTeX formatting
kable(summary_table, caption = "Summary Statistics", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>%
  row_spec(0, extra_latex_after = "\\hline\\hline") %>%  # Two lines at the top
  row_spec(nrow(summary_table), extra_latex_after = "\\hline")  # One line at the bottom
```